Every robotic system has some set of parameters---scale factors, sensor locations, link lengths, etc.---that are needed for state estimation, planning, and control. Despite best efforts during construction, some parameters will change over the lifetime of a robot due to normal wear and tear. In the best case, incorrect parameter values degrade performance. In the worst case, they cause critical safety issues.
% Calibration is important for performance of robotic systems.
%% Accuracy.
%% Safety.
We are interested in developing automated systems that are capable of robust long-term deployment in the hands of non-experts, so the identification and update of these parameter values is highly important.

As an example, consider a camera-based collision avoidance system intended for deployment in a consumer automobile. For the vehicle to be able to avoid collisions, the pose of each camera with respect to the vehicle coordinate system must be known precisely so that obstacle positions can be transformed from camera coordinates into vehicle coordinates. However, a consumer vehicle will have no access to special calibration hardware or expert data analysis. 
In such a scenario, the vehicle must be capable of self-supervised recalibration. This problem is inherently difficult for a number of reasons that we will briefly discuss here.

 {\bf 1) Parameters change over time}---although the vehicle may be factory calibrated, the transformations can change slowly over time due to vibration, thermal expansion, loose parts, or any number of other common problems that follow from normal usage. 

 {\bf 2) Parameters must be inferred from the data}---as the cameras may be installed in different places on the vehicle, their pose cannot be measured directly. Instead, the transformations must be inferred from the data produced by the full system. However, this is only possible if the motion of the vehicle renders the parameters {\em observable}\footnote{Informally, observability means that the parameters can be inferred from some local batch of measurement data \cite{jazwinski1970stochastic}}. 

 {\bf 3) Normal operation may result in unobservable directions in parameter space}---unfortunately, for this and many other practical problems normal operation may not render all directions in parameters space observable. In this example, when two cameras do not share an overlapping field of view, planar motion renders the calibration problem degenerate\footnote{See \cite{lebraly2010calibration} for a handy table of degenerate cases or \cite{kim2006absolute} for a more theoretical analysis.}. In fact, the transformation between cameras only becomes observable under general 3D motion. 

{\bf 4) Unobservable directions in parameter space may appear observable in the presence of noise}---even if our hypothetical car is piloted only on a plane (a degenerate case), noise in the measurements can make unobservable parameters appear observable. We call these parameters {\em numerically unobservable} to mirror the related concept of numerically rank-deficient matrices. 
%In a Kalman filter, parameter estimates take a random walk in the unobservable space and the estimator covariance will deterministically decrease. A batch estimator will jump directly to the wrong answer that minimzes the cost function. 

Existing algorithms for self calibration generally handle issues (1) and (2). Issue (3) is dealt with by designing experiments that guarantee all parameters become observable. To the best of our knowledge, no published self-calibration algorithm is able to cope with issue (4). Therefore, unless we plan to require all vehicle owners to outfit their parking places with calibration patterns or regularly drive off-road, new advances for online system calibration are required.


% What is hard:
%% Parameters change during normal usage
%% Parameters may not be directly measureable---they must be inferred from the data.
%% Some directions in parameter space may be unobservable (their values can not be determined from the measurements, cite Jazwinski) *unless the platform undergoes specific motions*
%%  * The concept of observability is deterministic but, in the presence of noise, parameters may appear observable even though they are not. We call this concept {\em numerically unobservable}

In this paper, we propose an algorithm to deal with {\em all} of these difficulties.
Our approach exploits the algebraic links between the Gauss-Newton algorithm, the Fischer Information Matrix, and nonlinear observability analysis to automatically detect directions in parameter space that are numerically unobservable and avoid updating our parameters in these directions; at any given time, parts of the parameter space that are observable will be updated based on the latest information, while unobservable parts will remain at their initial guess. Novel sets of measurements are detected using a mutual information test, and added to a working set that is used to estimate the parameters. The result is an algorithm that listens to an incoming stream of data, automatically accumulating a batch of data that may be used to calibrate an full robot system. The only requirement is that it is possible to implement a batch Gauss-Newton estimator for the system state and calibration parameters based on a set of measurements.



The remainder of the paper is structured as follows. Section~\ref{sec:rel}
will give a brief overview over related approaches. Section~\ref{sec:model} is
dedicated to the mathematical grounding of our method.
Section~\ref{sec:exp} demonstrates the validity of our approach through
extensive experiments and evaluation. Section~\ref{sec:conc} will conclude the
paper.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../main"
%%% End: 